\section{TODO}
%
%
\subsection{General}
%
%
%
\begin{itemize}
\item {\bf SPLIT THE APPLICATION INTO LIBRARIES. THE EXECUTABLE HAS MORE THAN 8 MB!!!}
\item Whenever an error is stated because a file was not found, 
then specify the entire path to the respective file.
\item This means that we have to properly treat relative and absolute paths.
\item Make sure that the \emph{asd}  scheme is found when SSR is started 
from a different working directory than \emph{SSR/bin}.
\item Either orientation or position should be sufficient as specification 
for plane waves in ASDF.
\item HOA
\item Implementation of directional sources
\item {\bf Dynamic scene handling}
\item InterSense client
\item More flexible assignment of virtual loudspeakers to physical outputs.
\item proper handling of moving sources (audible artefacts!!!)
\item optimize the convolver in terms of efficiency
\end{itemize}
%
%
\subsection{GUI}
%
\begin{itemize}
\item shadows
\item color gradients
\item interpolated textures
\end{itemize}
%
%
\subsection{WFS Renderer}
%
\begin{itemize}
\item Implementation of the 3dB/oct prefilter.
\end{itemize}
%
%
%
\subsection{VBAP Renderer}
%
\begin{itemize}
\item compensate for imperfect circular loudspeaker layouts via delay???
\end{itemize}
%
%
\subsection{Network Interface}
%
\begin{itemize}
\item Remote file dialog (to open files on the server from then client)
\end{itemize}
%
%
%
\subsection{(possible) bugs}
%
%
\begin{itemize}
\item There might be a bug in the loudspeaker selection for WFS. The setup file for
the linear array in flores works. But if you change the loudspeaker orientation from 180
to -180 degrees it won't work.
\end{itemize}
%